{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or, How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-09-01 thesis/landslide.tex
relwork relaxed mem checkers
icb complexity discussion

@@

Ultimately, the second half of the state space is pruned symmetrically.
In general, the number of ways to interleave two threads executing $N$ and $M$ events each is given by ${N+M \choose N}$;
in this case, the original state space's size was ${3+3 \choose 3} = 20$.
DPOR's reduction is characterized by replacing $N$ and $M$ with the number of {\em conflicting} events only;
in this case, ignoring all {\tt tmpN++} reorderings and testing only ${2+2 \choose 2} = 6$ branches.

@@

\begin{algorithm}[h]
	\For{$n \in [0..\mathsf{max}_P(U)]$}{
		\ForEach{$S \in U, P(S) \le n$}{
			Execute($S$)
		}
	}
	\caption{ICB exploration ordering}
	\label{alg:icb}
\end{algorithm}

%% stupid imperative way to write it
%\begin{algorithm}[h]
%	\For{$n \in [0..\infty]$}{
%		$\mathsf{skipped\_any} := \mathsf{false}$ \\
%		\ForEach{$S \in U$}{
%			\uIf{$P(S) \le n$}{
%				Execute($S$)
%			} \Else {
%				$\mathsf{skipped\_any} := \mathsf{true}$
%			}
%		}
%		\If{$\neg \mathsf{skipped\_any}$}{
%			{\bf Break}
%		}
%	}
%\end{algorithm}

@@

Therefore, Landslide's ICB implementation combines with DPOR
when tagging new branches to explore at the end of each branch:
just as DPOR skips alternate interleavings that are memory-independent,
ICB further filters interleavings requiring more preemptions than the current bound
out of the to-explore set.
The macro {\tt ICB\_BLOCKED},
defined in {\tt schedule.h},
decides if a given thread would require a preemption beyond the current bound to switch to%
\footnote{Since ``voluntary'' context switches (e.g. arising from {\tt yield()})
are often necessary for correct execution,
{\tt ICB\_BLOCKED} does not count such switches towards the preemption count.
Therefore, within a certain preemption bound $b$,
interleavings with more than $b$ context switches may still be tested.
}.
The DPOR implementation then checks, for some $I_{ij}$ it wants to mark for exploration,
whether {\tt ICB\_BLOCKED(}$T_j${\tt )} at the state after $t_i$,
and skips it if so ({\tt tag\_good\_sibling()}/{\tt tag\_all\_siblings()}).

@@

>>>>

\subsubsection{Example}
\label{sec:landslide-dpor-example}

@@

Ultimately, the second half of the state space is pruned symmetrically.
In general, the number of ways to interleave two threads executing $N$ and $M$ events each is given by ${N+M \choose N}$%
\footnote{Generalizing to $K$ threads, and simplifying to the same number $N$ of events each,
this formula becomes $\frac{n!k!}{n!^k}$.};
in this case, the original state space's size was ${3+3 \choose 3} = 20$.
DPOR's reduction is characterized by replacing $N$ and $M$ with the number of {\em conflicting} events only;
in this case, ignoring all {\tt tmpN++} reorderings and testing only ${2+2 \choose 2} = 6$ branches.

@@

\begin{algorithm}[h]
	\For{$B \in [0..\mathsf{max}_P(U)]$}{
		\ForEach{$S \in U, P(S) \le B$}{
			Execute($S$)
		}
	}
	\caption{ICB exploration ordering}
	\label{alg:icb}
\end{algorithm}

%% stupid imperative way to write it
%\begin{algorithm}[h]
%	\For{$B \in [0..\infty]$}{
%		$\mathsf{skipped\_any} := \mathsf{false}$ \\
%		\ForEach{$S \in U$}{
%			\uIf{$P(S) \le B$}{
%				Execute($S$)
%			} \Else {
%				$\mathsf{skipped\_any} := \mathsf{true}$
%			}
%		}
%		\If{$\neg \mathsf{skipped\_any}$}{
%			{\bf Break}
%		}
%	}
%\end{algorithm}

@@

Therefore, Landslide's ICB implementation combines with DPOR
when tagging new branches to explore at the end of each branch:
just as DPOR skips alternate interleavings that are memory-independent,
ICB further filters interleavings requiring more preemptions than the current bound
out of the to-explore set.
The macro {\tt ICB\_BLOCKED},
defined in {\tt schedule.h},
decides if a given thread would require a preemption beyond the current bound to switch to%
\footnote{Since ``voluntary'' context switches (e.g. arising from {\tt yield()})
are often necessary for correct execution,
{\tt ICB\_BLOCKED} does not count such switches towards the preemption count.
Therefore, within a certain preemption bound $B$,
interleavings with more than $B$ context switches may still be tested.
}.
The DPOR implementation then checks, for some $I_{ij}$ it wants to mark for exploration,
whether {\tt ICB\_BLOCKED(}$T_j${\tt )} at the state after $t_i$,
and skips it if so ({\tt tag\_good\_sibling()}/{\tt tag\_all\_siblings()}).

@@

\subsubsection{Complexity}

If the search is terminated early after reaching a predetermined fixed bound for $B$,
ICB in principle reduces the state space from exponentially-sized%
\footnote{Combinatorial, to be precise; see \sect{\ref{sec:landslide-dpor-example}}.}
in both $K$, the number of threads,
and $N$, the number of events,
to still exponential in $K$ (typically small) but only polynomial in $N$ (typically large).
Under a preemption bound of $B$, there are only $B+K$ opportunities for context switching%
\footnote{This $K$ appears from the ``mandatory'' context switches at thread exit;
more of which could also be introduced from blocking synchronization.},
% "roughly" - i.e., this formulation allows choosing all B preemptions from the same thread,
% which is only possible when K >= B i guess
% it's an overapproximation but getting it more precise seems like it'd be way more complex,
% and the overapproximation is good enough to make this point anyway
% also an underapprox because this doesn't count interleavings with b<B preemptions but that's a constant factor
so the corresponding state space size is at most ${KN \choose B}(B+K)!$.
All $N$-related factors therein are bounded above by $N^B$.

Prior work often recommends 2 for such a cutoff \cite{chess-icb,smc-empirical-study,dejafu},
although \sect{\ref{sec:quicksand-eval}}'s larger dataset suggests 3 would be considerably more thorough.
On the other hand, any finite such bound can provide only a heuristic verification guarantee anymore.
Preserving the full formal verification, i.e., continuing iteration until $B = \mathsf{max}_P(U)$
not only remains exponential in $N$,
% again overapproximating because each previous iteration is combinatorially smaller than the next
but also introduces a factor of $\mathsf{max}_P(U)$ repeated work.
Future work could memoize already-tested interleavings so that each iteration of $B$ could test only
those schedules with exactly $B$ preemptions,
restoring the original $B$-independent (but still exponential) complexity.
